TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2019-06-11 paper/draft.tex
eliminate redundancy in intro

Here we take M-dwarf labels from samples of well-characterized stars that overlap the
\sdssiv\ \apogee\ sample, and use those labels to train a model and label all of the M dwarfs observed by
\sdssiv\ \apogee. One set of labels are physical parameters (effective temperatures and metallicities)
and the other set of labels are spectral types. In each case, \thecannon\ produces a predictive model for \apogee\ spectra, but without
any input of stellar physics beyond whatever went in to producing the labels in the first
place. That is, the model to be applied to the \apogee\ data is learned from the \apogee\ data alone.
This obviates physical models that are often inaccurate for M dwarfs, and capitalizes on the excellent data quality
delivered by the \apogee\ spectrograph and project.

@@

The \apogee\ survey is a high resolution (R$\sim$22,500), H-band (1.5-1.7$\mu$m), multi-epoch survey which has observed over 250,000 stellar spectra up to its fourteenth data release (DR14; \citealt{Abolfathi:2017}). Fundamental parameters for each of these stars are estimated by the \apogee\ Stellar Parameter and Chemical Abundances Pipeline (\aspcap; \citealt{Perez:2016}), which employs a $\chi^2$ fitting procedure using the \textsl{FERRE} code to fit radiative transfer models and determine atmospheric parameters, 15 chemical abundances and micro-turbulence parameters ($\teff$, $\logg$, $\feh$, [$\alpha$/M], [C/M], [N/M]; \citealt{Meszaros:2012}). For low temperatures ($2800<\teff<3500$\,K), the pipeline uses \textsl{MARCS} plane-parallel/spherical models \citep{Gustafsson:2008}, and for higher temperatures ($\teff\geq3500$\,K) \textsl{ATLAS9} \citep{Castelli:2004} plane-parallel models are used.

>>>>

Here we take M-dwarf labels from samples of well-characterized stars that overlap the
\sdssiv\ \apogee\ sample, and use those labels to train a model and label all of the M dwarfs observed by
\sdssiv\ \apogee. One set of labels are physical parameters (effective temperatures and metallicities)
and the other set of labels are spectral types.

% In each case, \thecannon\ produces a predictive model for \apogee\ spectra, but without any input of stellar physics beyond whatever went in to producing the labels in the first place. That is, the model to be applied to the \apogee\ data is learned from the \apogee\ data alone. This obviates physical models that are often inaccurate for M dwarfs, and capitalizes on the excellent data quality delivered by the \apogee\ spectrograph and project. 
%JB: I think this part is redundant

@@

The \apogee\ survey is a high resolution (R$\sim$22,500), H-band (1.5-1.7$\mu$m), multi-epoch survey which has observed over 250,000 stellar spectra up to its fourteenth data release (DR14; \citealt{Abolfathi:2017}). Fundamental parameters for each of these stars are estimated by the \apogee\ Stellar Parameter and Chemical Abundances Pipeline (ASPCAP; \citealt{Perez:2016}), which employs a $\chi^2$ fitting procedure using the \textsl{FERRE} code to fit radiative transfer models and determine atmospheric parameters, 15 chemical abundances and micro-turbulence parameters ($\teff$, $\logg$, $\feh$, [$\alpha$/M], [C/M], [N/M]; \citealt{Meszaros:2012}). For low temperatures ($2800<\teff<3500$\,K), the pipeline uses \textsl{MARCS} plane-parallel/spherical models \citep{Gustafsson:2008}, and for higher temperatures ($\teff\geq3500$\,K) \textsl{ATLAS9} \citep{Castelli:2004} plane-parallel models are used.
